opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-23 chapters/04_our_dataset.tex
Updated figurre/texxt formatting in our_dataset

Here, PCA was performed on datasets of X-ray diffraction patterns. These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=N_{\text{values}} =512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines. Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N_{\text{values}}=512$. However, the maximal number of components is even lower for datasets that contain less than 512 patterns. In this case, the maximal number of components is equal to the number of patterns in the dataset since each pattern can add at most one degree of freedom to the dataset $X \in \mathbb{R}^N$. Hence the maximum number of components $N_{\text{max}}$ of a pattern dataset $X$ is given as follows:
\begin{align}
    N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
\end{align}
Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$.

\pagebreak

The below figure~$\eqref{fig:components}$ shows the cumulative explained variance ratio over the fraction of maximal No. components $N_{\text{max}}$ as defined above. In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\linewidth]{figures/component_fractions.png}
    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD. Each dataset is named after the contributing institutions. Datasets contributed by the same institution are enumerated alphabetically in the order in which they are described below. }
    \label{fig:components}
\end{figure}

%TODO: A paragraph on how data is correlated
The degree of variation between the patterns is different for each contribution. For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes. \\

\pagebreak

>>>>

Here, PCA was performed on datasets of X-ray diffraction patterns. These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=N_{\text{values}} =512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines. Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N_{\text{values}}=512$. However, the maximal number of components is even lower for datasets that contain less than 512 patterns. In this case, the maximal number of components is equal to the number of patterns in the dataset since each pattern can add at most one degree of freedom to the dataset $X \in \mathbb{R}^N$. Hence the maximum number of components $N_{\text{max}}$ of a pattern dataset $X$ is given as follows:
\begin{align}
    N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
\end{align}
Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$. Figure~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No. components $N_{\text{max}}$ as defined above. In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar. \\

The degree of variation between the patterns is different for each contribution. For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.  
\pagebreak

These effects can be observed in figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by the USC and LBNL converge towards one significantly faster than the datasets contributed by the CNRS and HKUST. \\

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD. Each dataset is named after the contributing institutions. Datasets contributed by the same institution are labeled alphabetically in the order in which they are described in the texts towards the end of this section.}
    \label{fig:components}
\end{figure}
